---
title: "Covid91 vaccine study Final2022"
author: "DESMOND KOUTSO"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final_Data.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
Females=filter(FinalData,sex=="F")
LGBTQ=filter(FinalData,LGBTQ=="gay")
LGBTQ=filter(FinalData,LGBTQ=="straight")
DrugUser=filter(FinalDat,DrugUser=="no")
DrugUser=filter(FinalDat,DrugUser=="yes")
```

# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

### Null and Alternative hypothesis

$H_O$: . There is no difference between how effective the vaccine is for any of the subgroup that uses it.These two variables are independent.    

$H_A$: There is a difference between how effective the vaccine is for any of the subgroup that uses it. These two variables are dependent.  

### Methods

This data set is a CAT~CAT since both variables are categorical. Whether the vaccine is effective or not can be categorized and also, the subgroups can be categorized. A barplot will be used to describe the graphical results, a data table for the numerical results, a chi-squared test and fisher exact test to describe the inferential results for the different 4 subgroups. To end I will give my recommendation on whether or not each sub-group should take the vaccine.   

## Males

### Graphical Descriptive results


```{r}
barchartGC(~infected + treatment,data=FinalData)
barchartGC(~infected + treatment,data=FinalData, type="percent")

```

On the second Plot which represents the percentage of male that took the vaccine to those who did not, it seen that the male that took the vaccine have a lower percentage(chance) of contracting the Covid19.  
On the first plot, the frequency of males that the vaccine is slightly lower to contract the covid19 as compared to those who took the placebo drug.  
Over all, the men that contracted covid19 the most are the ones that took the placebo drug, and so the vaccine is effective in males for previnting the covid19.  

### Numerical Descriptive results


```{r}
table1 <- xtabs(~infected + treatment,data=FinalData)
rowPerc(table1)
colPerc(table1)
```
The table shows that if a  male  was to get the vaccine, he will be more likely to not contract the virus.  

### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
ftestM=fisher.test(table1)
ftestM
effecM=ftestM$estimate/(ftestM$estimate+1)*100
```

Based on the inferential results, the chi-squared test gives a p-value of 0.00000000000012, which is less than 0.05, so we reject the null hypothesis. 
The fisher's exact test gives a P-value of 0.00000000000019.  
The odds ratio tells us that the males that take the placebo are 1.7 times more likely to contract covid19 than those that took the vaccine .    
the efficacy is `r effecM`%

### Male Conclusion

For the male, it is observed from the analysis and from the efficacy that the vaccine is effective in preventing them from getting the virus. For this reason, I recommend the males to take the vaccine.  

## Females

### Graphical Descriptive Results

```{r}
barchartGC(~infected + treatment,data=Females)
barchartGC(~infected + treatment,data=Females, type="percent")

```

Based on this plot the females that took the placebo have a greater risk of contracting the virus.     

### Numerical Descriptive Results

```{r}
table1 <- xtabs(~infected + treatment,data=Females)
rowPerc(table1)
colPerc(table1)
```

the table shows that the females that took the placebo are more likely to contract the virus.


### Inferential results

```{r}
chisq.test(table1)
chisqtestGC(table1)
ftestF=fisher.test(table1)
ftestF
effecF=ftestF$estimate/(ftestF$estimate+1)*100

```

from the chi-squared test, the p-value is 0.0011 which is less than 0.05, so we reject the Null hypothesis.  
From the fisher's exact test, the p-value is 0.00094.  
the odds ratio shows that the females that take the placebo are 1.4 times likely to contract the covid19 than those that took the vaccine .   
the efficacy is `r effecF`%  

### Females Conclusion
to conclude, since from the analysis and the drug efficacy(58.8%)it is seen that the females that take the vaccine are more likely to not contract the the virus,so I recommend that the subgroup females should use the vaccine.    

## LGBTQ

### Graphical Descriptive Results

```{r}
barchartGC(~infected + treatment,data=LGBTQ)
barchartGC(~infected + treatment,data=LGBTQ, type="percent")
```

From this plot, in the subgroup LGBTQ, the people that took the placebo have a greater percentage of contracting the covid19 as compared to those that took the vaccine, they have a lower percentage of contracting covid19.  

### Numerical Descriptive Results

```{r}
table1 <- xtabs(~infected + treatment,data=LGBTQ)
rowPerc(table1)
colPerc(table1)
```

The table shows that, in the LGBTQ subgroup, those that took the placebo have a greater rate of getting covid19(67.88) while those that took the vaccine have lower rate of getting covid19(32.12).  

### Inferential results

```{r}
chisq.test(table1)
chisqtestGC(table1)
ftestLGBTQ=fisher.test(table1)
ftestLGBTQ
effecLGBTQ=ftestLGBTQ$estimate/(ftestLGBTQ$estimate+1)*100
```

The chi-squared test give a p-value of 0.0011.  
The fisher's exact test gives a p-value of 0.0009484.  
Both the p-value of the fisher's exact test and the chi-squared test causes us to reject the null hypothesis because they are less than 0.05.  
The odds ratio tells us that the LGBTQ subgroup that took the placebo are 1.43 times more likely to contract covid19 than those that took the vaccine.       
the efficacy is `r effecLGBTQ`%   

### LGBTQ Conclusion

To conclude, due to the analysis, and the efficacy of  the vaccine on the LGBTQ subgroup(58.8), I therefore recommend that that the vaccine should be used on the LGBTQ subgroup.  

## Druggies

### Graphical Descriptive Results

```{r}
barchartGC(~infected + treatment,data=DrugUser)
barchartGC(~infected + treatment,data=DrugUser, type="percent")

```

from this plot, among the drug users,it is observed that taking the vaccine puts them a geater risk of getting the covid19.   

### Numerical Descriptive Results

```{r}
table1 <- xtabs(~infected + treatment,data=DrugUser)
rowPerc(table1)
colPerc(table1)
```

From this table, the drug users that took the vaccine have a greater risk of getting covid19(85.00).      

### Inferential results

```{r}
chisq.test(table1)
chisqtestGC(table1)
ftestDrugUser=fisher.test(table1)
ftestDrugUser
effecDrugUser=ftestDrugUser$estimate/(ftestDrugUser$estimate+1)*100
```

The chi-squared test gives a p-value of 0, which is  less than 0.05,so we reject the null hypothesis.    
On the fisher's exact test, the odds ratio says that the drugusers that took the vaccine are 0.017 more likely to get covid19 than those that took the placebo  .     
the efficacy is `r effecDrugUser`%     

### Druggies Conclusion

the analysis show that the vaccine has no effect on the drug users as the ones that took the vaccine had the greatest rate of contracting the virus. More to that, the efficacy of the drug is 15.1, meaning it is not  effective on the subgroup DrugUsers, and so, I do not recommend the vaccine on druguser.     

## Final Conclusion

The question asked was if the Covid-19 vaccine exhibits different effectiveness across the variety of subgroups.  
Based on the general Null and Alternative hypothesis, I reject the null hypothesis because there was eventually different effectiveness of the vaccine across the different subgroups.  
The vaccine was effective on the males, and it was eventually effective on the females,and so I recommended them to use the Vaccine. The vaccine was equally effective on the  LGBTQ subgroup and so i recommended that they should use the vaccine. The  Druggies had no effectiveness of the vaccine on them, and so i did not recommend the vaccine to them.     