---
title: "Covid91 vaccine study Final2022"
author: "DESMOND KOUTSO"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final_Data.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

### Null and Alternative hypothesis

$H_O$: There is no difference between how effective the vaccine is for any of the sub-groups that uses it. These two variables are independent.  

$H_A$: There is a difference between how effective the vaccine is for any of the subgroup that uses it. These two variables are dependent.  

### Methods

This data set is a CAT~CAT since both variables are categorical. Whether the vaccine is effective or not can be categorized and also, the subgroups can be categorized. A barplot will be used to describe the graphical results, a data table for the numerical results, a chi-squared test and fisher exact test to describe the inferential results for the different 4 subgroups. To end I will give my recommendation on whether or not each sub-group should take the vaccine.   

## Males

### Graphical Descriptive results


```{r}
barchartGC(~infected + treatment,data=FinalData)
barchartGC(~infected + treatment,data=FinalData, type="percent")

```

On the first Plot which represents the percentage of male that took the vaccine to those who did not, it seen that the male that took the vaccine have a lower percentage(chance) of contracting the Covid19.  
On the second plot, the frequency of males that the vaccine is slightly lower to contract the covid19 as compared to those who took the placebo drug.  
Over all, the men that contracted covid19 the most are the ones that took the placebo drug, and so the vaccine if effective in males for previnting the covid19.  

### Numerical Descriptive results


```{r}
table1 <- xtabs(~infected + treatment,data=FinalData)
rowPerc(table1)
colPerc(table1)
```
The table shows that if a  male  was to get the vaccine, he will be more likely to not contract the virus.  

### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

Based on the inferential results, the chi-squared test gives a p-value of 0.00000000000012, which is less than 0.05, so we reject the null hypothesis. 
The fisher's exact test gives a P-value of 0.00000000000019.  
The odds ratio tells us that the males that take the vaccine hvae 1.7 chances to not contract the virus as compared to those that took the placebo .  

### Male Conclusion

For the male, it is observed from the analysis that the vaccine is effective in preventing them from getting the virus. For this reason, I recommend the males to take the vaccine.  

## Females

### Graphical Descriptive Results

### Numerical Descriptive Results

### Inferential results

### Females Conclusion

## LGBTQ

### Graphical Descriptive Results

### Numerical Descriptive Results

### Inferential results

### LGBTQ Conclusion


## Druggies

### Graphical Descriptive Results

### Numerical Descriptive Results

### Inferential results

### Druggies Conclusion

## Final Conclusion
