---
title: "Covid91 vaccine study Final2022"
author: "DESMOND KOUTSO"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final_Data.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
Females=filter(FinalData,sex=="F")
```

# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

### Null and Alternative hypothesis

$H_O$: . There is no difference between how effective the vaccine is for any of the subgroup that uses it.These two variables are independent.    

$H_A$: There is a difference between how effective the vaccine is for any of the subgroup that uses it. These two variables are dependent.  

### Methods

This data set is a CAT~CAT since both variables are categorical. Whether the vaccine is effective or not can be categorized and also, the subgroups can be categorized. A barplot will be used to describe the graphical results, a data table for the numerical results, a chi-squared test and fisher exact test to describe the inferential results for the different 4 subgroups. To end I will give my recommendation on whether or not each sub-group should take the vaccine.   

## Males

### Graphical Descriptive results


```{r}
barchartGC(~infected + treatment,data=FinalData)
barchartGC(~infected + treatment,data=FinalData, type="percent")

```

On the second Plot which represents the percentage of male that took the vaccine to those who did not, it seen that the male that took the vaccine have a lower percentage(chance) of contracting the Covid19.  
On the first plot, the frequency of males that the vaccine is slightly lower to contract the covid19 as compared to those who took the placebo drug.  
Over all, the men that contracted covid19 the most are the ones that took the placebo drug, and so the vaccine is effective in males for previnting the covid19.  

### Numerical Descriptive results


```{r}
table1 <- xtabs(~infected + treatment,data=FinalData)
rowPerc(table1)
colPerc(table1)
```
The table shows that if a  male  was to get the vaccine, he will be more likely to not contract the virus.  

### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

Based on the inferential results, the chi-squared test gives a p-value of 0.00000000000012, which is less than 0.05, so we reject the null hypothesis. 
The fisher's exact test gives a P-value of 0.00000000000019.  
The odds ratio tells us that the males that take the vaccine hvae 1.7 chances to not contract the virus as compared to those that took the placebo .  

### Male Conclusion

For the male, it is observed from the analysis that the vaccine is effective in preventing them from getting the virus. For this reason, I recommend the males to take the vaccine.  

## Females

### Graphical Descriptive Results

```{r}
barchartGC(~infected + treatment,data=Females)
barchartGC(~infected + treatment,data=Females, type="percent")

```
Based on this plot the females that took the placebo have a greater risk of contracting the virus.     

### Numerical Descriptive Results

```{r}
table1 <- xtabs(~infected + treatment,data=Females)
rowPerc(table1)
colPerc(table1)
```

the table shows that the females that took the placebo are more likely to contract the virus.  
### Inferential results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

from the chi-squared test, the p-value is 0.0011 which is less than 0.05, so we reject the Null hypothesis.  
From the fisher's exact test, the p-value is 0.00094.  
the odds ratio shows that there are 1.4 chances that the females that take the vaccine will not contract the vaccine. 


### Females Conclusion
to conclude, since from the analysis it is seen that the females 
that take the vaccine are more likely to not contract the the vaccine,so I recommend that the subgroup females should use the vaccine.  

## LGBTQ

### Graphical Descriptive Results

```{r}
barchartGC(~treatment + LGBTQ,data=FinalData)

```

From this plot, after the gay and straight take the placebo medication and the vaccine, the frequency of the effect of gay that took the vaccine and straight that took the vaccine is the same.  
the straight that took the vaccine and the placebo too, have the same frquency on the effectivenes of the placebo and the vaccine.

### Numerical Descriptive Results

```{r}
table1 <- xtabs(~treatment + LGBTQ,data=FinalData)
rowPerc(table1)
colPerc(table1)
```

The table shows that, after the straight and the gay have received the vaccine and placebo, there is a slight(little) or no different of the effect of the vaccine or placebo on the gays and the straight.  

### Inferential results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

From the table above, the chi-squared test gives us a p-value of 0.8924, so we fail to reject the null hypothesis by saying that There is no difference between how effective the vaccine is for either the gays or the straight in LGBTQ.    
From the exact test, the P-value is 0.88, which is greater than 0.05 and so we fail to reject the null hypothesis.  
The odds ratio tells us that there are 1.005 chances that the vaccine or the placebo have no effect on the gays and straight in the subgroup LBTQ.

### LGBTQ Conclusion

To conclude, there is no difference in the effect of vaccine or the placebo for the gays and the straight in LGBTQ subgroup. For this reason, i recommend that the the subgroup LGBTQ should not receive the vaccine.  

## Druggies

### Graphical Descriptive Results

```{r}
barchartGC(~treatment + DrugUser,data=FinalData)

```

from this table, there is no difference in the effect of either the placebo or the vaccine on those who use drugs or no.  

### Numerical Descriptive Results

```{r}
table1 <- xtabs(~treatment + DrugUser,data=FinalData)
rowPerc(table1)
colPerc(table1)
```

From this table, there is a slight(little) difference or no in the effect of the placebo or the vaccine on either those that use drugs or not.  

### Inferential results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The chi-squared test gives a p-value of 0.18, which is greater than 0.05,so we fail to reject the null hypothesis.  
On the fisher's exact test, the odds ratio says that there is 1.05 chance that the vaccine or placebo have no effect on either the drug users or those that do not use drugs.   

### Druggies Conclusion

Since the there are no difference on effectiveness of the placebo or the vaccine on the drug users or the non drug users, I do not recommend the vaccine for the subgroup Druggies.

## Final Conclusion

The question asked was if the Covid-19 vaccine exhibits different effectiveness across the variety of subgroups.  
Based on the general Null and Alternative hypothesis, I reject the null hypothesis because there was eventually different effectiveness of the vaccine across the different subgroups.  
The vaccine was effective on the males, and it was eventually effective on the females,and so I recommended them to use the Vaccine. The LGBTQ subgroup and the Druggies had no effectiveness of the vaccine on them, and so i did not recommend the vaccine to them.   